---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(width = 110)

library(fetchdhs)
library(tidyverse)
```

[![Travis-CI Build Status](https://travis-ci.org/murphy-xq/fetchdhs.svg?branch=master)](https://travis-ci.org/murphy-xq/fetchdhs)

# fetchdhs

A suite of convenience functions to retrieve tidy dataframes of survey data from the [DHS API](https://api.dhsprogram.com/)

## Installation
```{r installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("murphy-xq/fetchdhs")
```

## Basic example

Let's say you quickly need DHS survey data for:

* immunization-related indicators
* since 2000
* for India and Nigeria

First, use `fetch_countries()` to locate the 2-letter DHS country codes for India and Nigeria needed for the api call

```{r country_list_filter}
fetch_countries() %>% 
  filter(country_name %in% c("India", "Nigeria"))
```

Next, make use of [DHS API tags](https://api.dhsprogram.com/rest/dhs/tags?f=html) that categorize survey indicators by topic. In this example, we are looking for all immunization-related indicators using `fetch_tags()` and identify tag `32`

```{r tags_list}
fetch_tags() %>% 
  filter(str_detect(tag_name,  "[Ii]mmunization"))
```

Finally, use `fetch_data()` to call the DHS API using the parameters just identified and receive a tidy dataframe as well as the api call:

* Note only 1 tag may be specified

```{r fetch_dta}
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017)
```

For specific indicators, we can peek at a dataframe of all available indicators to identify which `indicator_id` codes should be included with `fetch_data()`. Let's try pulling only DPT3 and Measles indicators:

```{r find_indicators}
fetch_indicators() %>% 
  filter(str_detect(definition, "Measles|DPT3"))
```

Upon investigating the DPT3 and Measles indicators and their associated [attributes](https://api.dhsprogram.com/rest/dhs/indicators/fields), we see that we need to use `CH_VACC_C_DP3` and `CH_VACC_C_MSL`:

```{r fetch_indicator}
fetch_data(countries = c("IA","NG"), indicators = c("CH_VACC_C_DP3", "CH_VACC_C_MSL"), years = 2000:2017)
```

## Additional features

#### Level of disaggregation

We have been using the default level of disaggregation which returns national-level data only. In order to pull subnational, background characteristic, or all available data, we need to specify the `breakdown_level` in `fetch_data()`

* Note the increasing amount of records returned as we move from the default `breakdown_level == "national"` (205 records) to `breakdown_level == "all"` (3,270 records)

```{r breakdown}
# national (default)
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017, breakdown_level = "national")

# subnational
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017, breakdown_level = "subnational")

# background
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017, breakdown_level = "background")

# all
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017, breakdown_level = "all")
```

#### Return fields

[Return fields](https://api.dhsprogram.com/rest/dhs/data/fields) are the various dimensions of survey data that can be returned. `set_return_fields()` allows the user to specify which fields should comprise the dataframe returned from the api

```{r return_field}
set_return_fields(c("Indicator", "CountryName", "SurveyYear", "SurveyType", "Value"))

fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017)
```

#### Add geospatial data

We can also include polygon coordinates with our call with `add_geometry`

```{r add_geom}
fetch_data(countries = c("IA","NG"), tag = 32, years = 2000:2017, add_geometry = TRUE)
```


#### API Key

Authenticated users can query more records per page -- 5,000 versus 1,000 maximum records per page. Please see here for [authentication details](https://api.dhsprogram.com/#/introdevelop.html).

Users can input their api key with `set_api_key()` for inclusion in any subsequent `fetch_data()` calls.

```{r set_api_key}
set_api_key("YOURKEY-GOESHERE")
```